Adaptive Approximation for Multivariate Linear Problems with Inputs Lying in a Cone

### 2019-09-14 DingHickKritzMakRICAMChapVerD.tex
Updates from Overleaf

    \RETURN $\ALG(f,\varepsilon) = \APP(f,n^*)$
\end{algorithmic}
\end{algorithm}
We now show that this algorithm is optimal if the input set is a ball.
\begin{theorem}\label{DHKM:thm_cost_non_adapt}
The non-adaptive Algorithm \ref{DHKM:BallAlg} has an information cost for the set of input functions $\calC = \calB_R$ that is given by
\[
\COST(\ALG, \calB_R, \varepsilon,R') = \min \left \{n \in \N_0 : \bignorm[\rho']{\bslambda^{(n+1)}} \le \varepsilon /R \right \}.
\]
This algorithm is essentially optimal for the set of input functions $\calB_R$, namely,
\begin{multline*}
\COST(\ALG, \calB_R, \varepsilon,R') \le \COMP(\calA(\calB_R),\omega \varepsilon ,R') \\ \forall \, \varepsilon \in (0, \varepsilon_{\max}], \ R \in [R_{\min}, \infty),
\end{multline*}
where $\varepsilon_{\max}$ and $R_{\min}$ are arbitrary and fixed, and $\omega = R_{\min}/R$.
\end{theorem}

@@

\YuhanNote{Create a new subsection and move the content related with the example here}
\DHKMchange{
We now derive upper and lower bounds on the information costs of Algorithms \ref{DHKM:BallAlg} and \ref{DHKM:PilotConeAlg} for the case  $\rho = \tau$ and $\rho' = \infty$, which includes the example introduced in Section \ref{DHKM:secexamp}.  For $\rho'=\infty$, $\bignorm[\rho']{\bslambda^{(n+1)}} = \lambda_{\bsk_{n+1}}$.  To facilitate these bounds we establish a couple of commonly used lemmas.

@@

\FredNote{As far as simplifications go,

1) If we like the $\bslambda^{(n)}$ notation and even want to write $\bslambda^{(n)}_d$ to emphasize the $d$ dependence, then we can simplify some of the equations below.}

\PeterNote{I am not sure. I have the impression that with the additional restriction 
that many sums do not start with index 1, but with indices depending on $d$ and other parameters, it might make notation overly complicated... .
I mostly left it as it is. But if you prefer, we can still change it!}

\FredNote{I commented out where I thought we could introduce the notation.  Let me know what you think.}

\PeterNote{I agree. Let us use your notation, it makes it more consistent.}

@@

\begin{theorem} \label{DHKM:thmtract1}
Using the same notation as above, the following statements hold for the case $\rho'=\infty$.
 \begin{itemize}
  \item[1.] 
  We have strong polynomial tractability if and only if there exist $p>0$ and $i_0\in\NN$ such that
%  \begin{equation}\label{DHKM:eq:condspt}
%     \sup_{d\in\NN} \sum_{i=i_0}^\infty \lambda_{d,\bsk_i}^p < \infty.
%  \end{equation}
 \begin{equation}\label{DHKM:eq:condspt}
    \sup_{d\in\NN} \, \bignorm[p]{\bslambda_d^{(i_0)}} < \infty.
 \end{equation}
 Furthermore, the exponent of strong polynomial tractability is then equal to the infimum of those $p>0$ for which \eqref{DHKM:eq:condspt} holds. 
 \item[2.] 
  We have polynomial tractability if and only if there exist $p_1, p_2 \ge 0$ and $p_3, K>0$ such that
%  \[
%     \sup_{d\in\NN} d^{-p_1}\, \sum_{i=\lceil K d^{p_2} \rceil}^\infty \lambda_{d,\bsk_i}^{p_3} < \infty.
%  \]
 \[
    \sup_{d\in\NN} d^{-p_1}\, \bignorm[p_3]{\bslambda_d^{(\lceil K d^{p_2} \rceil)}} < \infty.
 \]
 \item[3.] 
 We have weak tractability if and only if 
 \begin{equation}\label{DHKM:eq:condwt}
  \sup_{d\in\NN} \, \exp(-cd) \sum_{i=1}^\infty \exp\left(-c\left(\frac{1}{\lambda_{d,\bsk_i}}\right)\right) <\infty\quad \mbox{for all}\quad c>0.
 \end{equation}
\end{itemize}
\end{theorem}

\begin{proof}
Letting $\widetilde{\varepsilon}:=\sqrt{\varepsilon/R}$, we see that 
$n_{\COMP}(\varepsilon/R,d)=\min \left \{n\ge 0\colon \lambda_{d,\bsk_{n+1}} \,\le \widetilde{\varepsilon}^{\,2} \right \}$. The latter expression is well studied in the context of tractability of linear problems in the worst-case setting defined on unit balls of certain spaces, and if and only if conditions on the $\lambda_{d,\bsk_i}$ for various tractability notions are known. These conditions can be found in \cite[Chapter 5]{NovWoz08a} for (strong) polynomial tractability and  \cite{WerWoz17} for weak tractability.

@@

In this case, letting $\widetilde{\varepsilon}:=(\varepsilon/R)^{\rho'/2}$ and 
$\widetilde{\lambda}_{d,i}:=\lambda_{d,\bsk_i}^{\rho'}$, we have 
% \begin{align}
%     \nonumber 
%     n_{\COMP}(\varepsilon/R,d) &=\min \left \{n\ge 0\colon 
% \sum_{i=n+1}^\infty \lambda_{d,\bsk_i}^{\rho'}\,
%     \le (\varepsilon/R)^{\rho'} \right \} \\
%     \label{DHKM:eqaveragetract}
%     & =\min \left \{n\ge 0\colon 
% \sum_{i=n+1}^\infty \widetilde{\lambda}_{d,i}\,
%     \le \widetilde{\varepsilon}^{\,2} \right \}.
% \end{align}
\begin{align}
    \nonumber 
    n_{\COMP}(\varepsilon/R,d) &=\min \left \{n\ge 0\colon 
\bignorm[\rho']{\bslambda_d^{(n+1)}}\,
    \le (\varepsilon/R)^{\rho'} \right \} \\
    \label{DHKM:eqaveragetract}
    & =\min \left \{n\ge 0\colon 
\bignorm[1]{\widetilde{\bslambda}_d^{(n+1)}}\,
    \le \widetilde{\varepsilon}^{\,2} \right \}.
\end{align}
However, the latter expression corresponds exactly to the 
average-case tractability (with respect to the parameters 
$\widetilde{\lambda}_{d,i}$ and $\widetilde{\varepsilon}$) defined 
on certain spaces as studied in, e.g., \cite{NovWoz08a}. 
This leads us to the following theorem.
\begin{theorem} \label{DHKM:thmtract2}
Using the same notation as above, the following statements hold for the case $\rho'<\infty$.
 \begin{itemize}
  \item[1.] 
  We have strong polynomial tractability if and only if there exist $p\in (0,1)$ and $i_0\in\NN$ such that
%  \begin{equation}\label{DHKM:eq:condspt1}
%     \sup_{d\in\NN} \sum_{i=i_0}^\infty \lambda_{d,\bsk_i}^{\rho'\,p} < \infty.
%  \end{equation}
 \begin{equation}\label{DHKM:eq:condspt1}
    \sup_{d\in\NN} \bignorm[\rho, p]{\bslambda_d^{(i_0)}} < \infty.
 \end{equation}
 Furthermore, the exponent of strong polynomial tractability is then 
 \[
 \inf\left\{\rho'p/(1-p)\colon \mbox{$p$ satisfies \eqref{DHKM:eq:condspt1}}\right\}.
 \]
 \item[2.] 
  We have polynomial tractability if and only if there exist $p_1, p_2 \ge 0$ and $p_3\in (0,1), K>0$ such that
%  \[
%     \sup_{d\in\NN} d^{-p_1}\, \sum_{i=\lceil K d^{p_2} \rceil}^\infty \lambda_{d,\bsk_i}^{\rho'\,p_3} < \infty.
%  \]
  \[
    \sup_{d\in\NN} d^{-p_1}\, \bignorm[\rho' p_3]{\bslambda_d^{(\lceil K d^{p_2} \rceil)}} < \infty.
 \]
 \item[3.] Let $t_{d,i}:=\sum_{\ell=i}^\infty \lambda_{d,\bsk_{\ell}}$. \FredNote{Here we have a problem since $i$ is both a summation index and limit.} \PeterNote{I think it should be ok now (?).}
 We have weak tractability if and only if 
 \[
   \lim_{i\to\infty} t_{d,i}\, (\log i)^2=0\quad\mbox{for all $d$},
 \]
 and there exists a function $f:[0,1/2)\to \{1,2,3,\ldots\}$ such that
\[
  \sup_{\beta\in (0,1/2]}\, \beta^{-2} \,
  \sup_{d\ge f(\beta)}\,\, \sup_{i\ge \lceil \exp (d\sqrt{b}) \rceil +1}\, \, \lim_{i\to\infty} t_{d,i}\, (\log i)^2
  < \infty.
\]
 \end{itemize}
\end{theorem}
\begin{proof}
  The proof of the theorem is similar to that of Theorem \ref{DHKM:thmtract1}, using \eqref{DHKM:eqaveragetract}.
\end{proof}
\begin{remark}
  Results for further tractability notions, such as quasi-polynomial tractability or $(s,t)$-weak 
  tractability, can be shown using similar arguments as above and results from \cite{KriWoz19}, \cite{NovWoz10a}, \cite{WerWoz17}, and the papers cited therein. 
\end{remark}

\DHKMchange{
To be more concrete, we return to the situation where the $\lambda_{d,\bsk}$ are given 
as in \eqref{DHKM:prodwts}  one more time. We note that
% Below, we write $\bslambda_d$ for the 
% sequence of the $\lambda_{d,\bsk}$
% Considering Condition \eqref{DHKM:eq:condspt} in Theorem \ref{DHKM:thmtract1}, note that
% \[
% \sup_{d\in\NN} \sum_{i=i_0}^\infty \lambda_{d,\bsk_i}^p = \sup_{d\in\NN} \norm[p]{\bslambda_d^{(i_0)}}^p.
% \]
% Since we assumed that the $\lambda_{d,\bsk}$ are ordered, 
$\bignorm[p]{\bslambda_d^{(i_0)}}^p<\infty$ holds if and only if  $\bignorm[p]{\bslambda_d}^p<\infty$.
We denote the $d$-dependence of the vector of coordinate weights by writing $\bsw_d:=(w_1, \ldots, w_d)$ and $\bsw_\infty = (w_1, w_2, \ldots)$
By Lemma \ref{DHKM:prodwtsupperbound}, we have
\[
\norm[p]{\bss} \norm[p]{\bsw_d} \le \norm[p]{\bslambda_d}\le \exp\left(\frac 1p \norm[p]{\bss}^p \norm[p]{\bsw_d}^p  \right),
\]
and this estimate holds for any $d\in\NN$.
%\[
%\sum_{i=1}^{\infty}\lambda_{d, \bsk_{i}}^{\eta} \le \exp\left(\sum_{k=1}^\infty s_k^\eta %\sum_{\ell = 1}^{d} w_\ell^{\eta} \right)\le \exp\left(\sum_{k =1 }^{\infty} s_k^\eta \sum_{\ell = 1}^{\infty} w_\ell^{\eta} \right). 
%\]
%Moreover, it also follows that for any fixed positive integer $i_0$, the sum of the $\lambda_{d,\bsk_i}^\eta$ is bounded below as
%\begin{align*}
%    \sup_{d \in \N} \, \sum_{i=i_0}^{\infty}\lambda_{d, \bsk_{i}}^{\eta} & \ge w_1^\eta \sum_{\DHKMchange{i=i_0}}^{\infty}s_{\DHKMchange{k_i}}^{\eta} \qquad \parbox{8cm}{considering only $\bsk_i$ of the form $(k, 0, 0, \ldots, 0)$, \\
%    and ordering $\bss$ so that $\DHKMchange{s_{k_1} \ge s_{k_2} \ge \cdots}$,} \\
%    \sup_{d \in \N} \, \sum_{i=i_0}^{\infty}\lambda_{d, \bsk_{i}}^{\eta} & \ge  s_{1}^{\eta} \sum_{i=i_0}^{\infty} w_{\ell_i}^\eta
    %\qquad %\text{for } d \ge i_0 
    %\\ &
%    \qquad \parbox{8cm}{considering only $\bsk_i$ of the form\\  $(0, \ldots, 0,1,0 , \ldots, 0)$,
%    where the non-zero\\ component is at the $\ell$-th   position with 
%    $i_0\le \ell \le d$,\\ and ordering $\bsw$  so that $w_{\ell_1} \ge w_{\ell_2} \ge \cdots$.}
%\end{align*}
Thus, we have necessary and sufficient conditions for strong tractability.
\begin{corollary} \label{DHKM:sptexample_cor}
For the $\lambda_{d,\bsk}$ of the form \eqref{DHKM:prodwts} we have strong polynomial tractability if and only if there exists $p>0$ such that
 \begin{equation*}
%\label{DHKM:sptexample}
    \norm[p]{\bss} < \infty \text{ and } 
    \norm[p]{\bsw_\infty} < \infty.
\end{equation*}
In particular, considering example in Section \ref{DHKM:secexamp}, where $s_{k} = k^{-r}$, we have strong polynomial tractability if and only if there exists $p> 0 $ such that
\begin{equation*}
%\label{DHKM:sptexample}
   r p > 1 \text{ and } 
    \norm[p]{\bsw_\infty} < \infty.
\end{equation*}
\end{corollary}

@@

Let $(n_j)_{j\ge 0}$ be a strictly increasing sequence of non-negative integers.  This sequence may increase geometrically or algebraically. Define the sets of wavenumbers analogously to \eqref{DHKM:KOnedef},
\begin{equation*}
   n_{-1}=0, \qquad \calK_j := \{\bsk_{n_{j-1}+1}, \ldots, \bsk_{n_j}\} \quad \text{for } j \in \N_0.
\end{equation*}
If $n_0 = 0$, then $\calK_0$ is empty.  For any $f \in \calF$, define the norms of subsets of series coefficients:
\begin{equation} \label{DHKM:SigmaDef}
\sigma_j (f):=\norm[\rho]{\biggl(\frac{\hf(\bsk)}{\lambda_{\bsk}} \biggr)_{\bsk \in \calK_j}} \qquad \text{for } j \in \N.
\end{equation}
Thus, $\norm[\calF]{f} = \norm[\rho]{\bigl(\sigma_j(f) \bigr)_{j \in \N_0}}$.

\DHKMchange{As stated above, we would like to keep track of the rate of decay of the $\hf(\bsk_i)$. Hence, for this section, we define the cone of input functions by}
\begin{equation} \label{DHKM:TrackConeDef}
  \DHKMchange{\calC_{\boldsymbol{\lambda},a,b}} : =\left\{f\in\calF \colon \sigma_{j+r} (f)\le ab^r \sigma_j (f)\ \forall j,r\in\NN\right\}.
\end{equation}
Here, $a$ and $b$ are positive reals with $b< 1 < a$. The constant $a$ is an inflation factor, and the constant $b$ defines the general rate of decay of the $\sigma_j(f)$ for $f \in  \DHKMchange{\calC_{\boldsymbol{\lambda},a,b}} $. 
\DHKMchange{We remark that knowing $\sigma_1 (f)$ would already be sufficient to identify 
a sample size that guarantees an error of at most $\varepsilon$. However, knowing $\sigma_j (f)$ for further values of $j$ will allow us to stop earlier.}

Because $ab^r$ may be greater than one, we do not require the series coefficients of the solution, $\SOL(f)$, to decay monotonically. However, we expect their partial sums to decay steadily.  The series coefficients for wavenumbers $\bsk \in \calK_0$ do not affect the definition of $\DHKMchange{\calC_{\boldsymbol{\lambda},a,b}} $ and may behave erratically.
Lemma \ref{DHKM:Key_Lem} implies that 
\begin{equation} \label{DHKM:LambdaDef}
    \norm[\tau]{\bigl(\hf(\bsk) \bigr)_{\bsk \in \calK_j}} \le \sigma_j(f) \Lambda_j, \qquad \text{where } \Lambda_j : = \norm[\rho']{\bigl(\lambda_{\bsk} \bigr)_{\bsk \in \calK_j}}.
\end{equation}
From \eqref{DHKM:SOLNorm} and \eqref{DHKM:SOLNormFinite} it follows that the norm of the solution operator is 
\begin{equation} \label{DHKM:NormLambdaFinite}
    \norm[\calF \to \calG]{\SOL} = \bignorm[\rho']{\bigl(\Lambda_j \bigr)_{j \in \N_0}} < \infty.
\end{equation}

If $f$ belongs to the $\DHKMchange{\calC_{\boldsymbol{\lambda},a,b}} $ defined in \eqref{DHKM:TrackConeDef} and $n_0 = 0$, then 
\begin{align*}
    \norm[\calF]{f} & = \norm[\rho]{\Biggl(\,\norm[\rho]{\biggl(\frac{\hf(\bsk)}{\lambda_{\bsk}} \biggr)_{\bsk \in \calK_j}}\,\Biggr)_{j \in \N}} = \bignorm[\rho]{\bigl(\sigma_j(f) \bigr)_{j \in \N}} \\
    & \le \bignorm[\rho]{\bigl(\sigma_1(f), ab \sigma_1(f), ab^2 \sigma_1(f), \ldots \bigr)} \\
    & = \left(1 + \frac{a^\rho b^\rho}{1 - b^\rho} \right)^{1/\rho}  \norm[\rho]{\biggl(\frac{\hf(\bsk)}{\lambda_{\bsk}} \biggr)_{\bsk \in \calK_1}}.
\end{align*}
Comparing this inequality to the definition of $\DHKMchange{\calC_{\boldsymbol{\lambda},n_1,A}} $ in the previous section, it can be seen that $\DHKMchange{\calC_{\boldsymbol{\lambda},a,b}} $  defined in \eqref{DHKM:TrackConeDef} is a subset of  $\DHKMchange{\calC_{\boldsymbol{\lambda},n_1,A}}$ defined in \eqref{DHKM:pilot_cone} if we choose 
$A=\left(1 + \frac{a^\rho b^\rho}{1 - b^\rho} \right)^{1/\rho}$ in \eqref{DHKM:pilot_cone}.

From the expression for the error in \eqref{DHKM:APP_Err_Coef} and the definition of the cone in  \eqref{DHKM:TrackConeDef}, we can now derive a data-driven error bound for all $f \in \DHKMchange{\calC_{\boldsymbol{\lambda},a,b}} $ and $j \in \bbN$: 
\begin{align}
\nonumber
\MoveEqLeft{\norm[\calG]{\SOL(f)-\APP(f,n_j)}} \\
\nonumber &= \norm[\tau]{\left(\hf(\bsk_i) \right)_{i = n_j+1}^\infty}
= \norm[\tau]{ \left(\norm[\tau]{\bigl(\hf(\bsk) \bigr)_{\bsk \in \calK_l}} \right)_{l=j+1}^\infty}
\\
\nonumber
& \le \norm[\tau]{ \bigl(\sigma_l(f) \Lambda_l \bigr)_{l=j+1}^\infty} \qquad \text{by \eqref{DHKM:LambdaDef}} \\
\nonumber 
&
= \norm[\tau]{ \bigl(\sigma_{j+r}(f) \Lambda_{j+r} \bigr)_{r=1}^\infty}
\\
& \le a \sigma_j(f) \norm[\tau]{ \bigl(b^r\Lambda_{j+r} \bigr)_{r=1}^\infty} =:\ERRNj
 \qquad \text{by \eqref{DHKM:TrackConeDef}.}
 \label{DHKM:algoineq}
\end{align}

@@

The data-driven error bound in \eqref{DHKM:algoineq} forms the basis for an adaptive Algorithm \ref{DHKM:TrackConeAlg}, which solves our problem for input functions in the cone $\DHKMchange{\calC_{\boldsymbol{\lambda},a,b}}$ defined in \eqref{DHKM:TrackConeDef}.  The following theorem establishes its viability and computational cost. In deriving upper bounds on the computational cost and lower bounds on the complexity, we may sacrifice tightness for simplicity.

@@

The following theorem establishes a lower bound on the complexity of our problem for input functions in $\DHKMchange{\calC_{\boldsymbol{\lambda},a,b}}$. The theorem after that shows that the cost of our algorithm as given in Theorem \ref{DHKM:TractConeCompCost} is essentially no worse than this lower bound.

>>>>

    \RETURN $\ALG(f,\varepsilon) = \APP(f,n^*)$
\end{algorithmic}
\end{algorithm}
We now show that this algorithm is optimal if the input set is a ball.
\begin{theorem}\label{DHKM:thm_cost_non_adapt}
The non-adaptive Algorithm \ref{DHKM:BallAlg} has an information cost for the set of input functions $\calC = \calB_R$ that is given by
\[
\COST(\ALG, \calB_R, \varepsilon,R') = \min \left \{n \in \N_0 : \bignorm[\rho']{\bslambda^{(n+1)}} \le \varepsilon /R \right \}, \qquad R’ \in (0, R].
\]
This algorithm is essentially optimal for the set of input functions $\calB_R$, namely,
\begin{multline*}
\COST(\ALG, \calB_R, \varepsilon,R') \le \COMP(\calA(\calB_R),\omega \varepsilon ,R') \\ \forall \, \varepsilon \in (0, \varepsilon_{\max}], \ R’ \in  [ R_{\min} , R], \ R \in [R_{\min}, \infty), 
\end{multline*}
where $\varepsilon_{\max}$ and $R_{\min}$ are arbitrary and fixed, and $\omega = R_{\min}/R$.
\end{theorem}

@@

\DHKMchange{
We now derive upper and lower bounds on the information costs of Algorithms \ref{DHKM:BallAlg} and \ref{DHKM:PilotConeAlg} for the case  $\rho = \tau$ and $\rho' = \infty$, which includes the example introduced in Section \ref{DHKM:secexamp}.  For $\rho'=\infty$, $\bignorm[\rho']{\bslambda^{(n+1)}} = \lambda_{\bsk_{n+1}}$.  To facilitate these bounds we establish a couple of commonly used lemmas.

@@

@@

\DHKMchange{\begin{theorem} \label{DHKM:thmtract1}
Using the same notation as above, the following statements hold for the case $\rho'=\infty$.
 \begin{itemize}
  \item[1.] 
  We have strong polynomial tractability if and only if there exist $p>0$ and $i_0\in\NN$ such that
 \begin{equation}\label{DHKM:eq:condspt}
    \sup_{d\in\NN} \, \bignorm[p]{\bslambda_d^{(i_0)}} < \infty.
 \end{equation}
 Furthermore, the exponent of strong polynomial tractability is then equal to the infimum of those $p>0$ for which \eqref{DHKM:eq:condspt} holds. 
 \item[2.] 
  We have polynomial tractability if and only if there exist $p_1, p_2 \ge 0$ and $p_3, K>0$ such that
 \[
    \sup_{d\in\NN} d^{-p_1}\, \bignorm[p_3]{\bslambda_d^{(\lceil K d^{p_2} \rceil)}} < \infty.
 \]
 \item[3.] 
 We have weak tractability if and only if 
 \begin{equation}\label{DHKM:eq:condwt}
  \sup_{d\in\NN} \, \exp(-cd) \sum_{i=1}^\infty \exp\left(-c\left(\frac{1}{\lambda_{d,\bsk_i}}\right)\right) <\infty\quad \mbox{for all}\quad c>0.
 \end{equation}
\end{itemize}
\end{theorem}
}

\begin{proof}
If $\widetilde{\varepsilon}:=\sqrt{\varepsilon/R}$, then
$n_{\COMP}(\varepsilon/R,d)=\min \left \{n\ge 0\colon \lambda_{d,\bsk_{n+1}} \,\le \widetilde{\varepsilon}^{\,2} \right \}$. The latter expression is well studied in the context of tractability of linear problems in the worst-case setting defined on unit balls of certain spaces, and if and only if conditions on the $\lambda_{d,\bsk_i}$ for various tractability notions are known. These conditions can be found in \cite[Chapter 5]{NovWoz08a} for (strong) polynomial tractability and  \cite{WerWoz17} for weak tractability.

@@

\DHKMchange{In this case, letting $\widetilde{\varepsilon}:=(\varepsilon/R)^{\rho'/2}$ and 
$\widetilde{\lambda}_{d,i}:=\lambda_{d,\bsk_i}^{\rho'}$, we have 
\begin{align}
    \nonumber 
    n_{\COMP}(\varepsilon/R,d) &=\min \left \{n\ge 0\colon 
\bignorm[\rho']{\bslambda_d^{(n+1)}}\,
    \le (\varepsilon/R)^{\rho'} \right \} \\
    \label{DHKM:eqaveragetract}
    & =\min \left \{n\ge 0\colon 
\bignorm[1]{\widetilde{\bslambda}_d^{(n+1)}}\,
    \le \widetilde{\varepsilon}^{\,2} \right \}.
\end{align}
However, the latter expression corresponds exactly to the 
average-case tractability (with respect to the parameters 
$\widetilde{\lambda}_{d,i}$ and $\widetilde{\varepsilon}$) defined 
on certain spaces as studied in, e.g., \cite{NovWoz08a}. 
This leads us to the following theorem.
\begin{theorem} \label{DHKM:thmtract2}
Using the same notation as above, the following statements hold for the case $\rho'<\infty$.
 \begin{itemize}
  \item[1.] 
  We have strong polynomial tractability if and only if there exist $p\in (0,1)$ and $i_0\in\NN$ such that
 \begin{equation}\label{DHKM:eq:condspt1}
    \sup_{d\in\NN} \bignorm[\rho, p]{\bslambda_d^{(i_0)}} < \infty.
 \end{equation}
 Furthermore, the exponent of strong polynomial tractability is then 
 \[
 \inf\left\{\rho'p/(1-p)\colon \mbox{$p$ satisfies \eqref{DHKM:eq:condspt1}}\right\}.
 \]
 \item[2.] 
  We have polynomial tractability if and only if there exist $p_1, p_2 \ge 0$ and $p_3\in (0,1), K>0$ such that
  \[
    \sup_{d\in\NN} d^{-p_1}\, \bignorm[\rho' p_3]{\bslambda_d^{(\lceil K d^{p_2} \rceil)}} < \infty.
 \]
 \item[3.] 
 We have weak tractability if and only if 
 \[
   \lim_{n\to\infty} (\log n)^2  \, \bignorm[1]{\bslambda_d^{(n)}} =0\quad\mbox{for all $d$},
 \]
 and there exists a function $f:[0,1/2)\to \{1,2,3,\ldots\}$ such that
\[
  \sup_{\beta\in (0,1/2]}\, \beta^{-2} \,
  \sup_{d\ge f(\beta)}\,\, \sup_{n \ge \lceil \exp (d\sqrt{b}) \rceil +1}\, \, (\log n)^2  \, \bignorm[1]{\bslambda_d^{(n)}} 
  < \infty.
\]
 \end{itemize}
\end{theorem}
}
\begin{proof}
  The proof of the theorem is similar to that of Theorem \ref{DHKM:thmtract1}, using \eqref{DHKM:eqaveragetract}.
\end{proof}
\begin{remark}
  Results for further tractability notions, such as quasi-polynomial tractability or $(s,t)$-weak 
  tractability, can be shown using similar arguments as above and results from \cite{KriWoz19}, \cite{NovWoz10a}, \cite{WerWoz17}, and the papers cited therein. 
\end{remark}

\DHKMchange{
To be more concrete, we return to the situation where the $\lambda_{d,\bsk}$ are given 
as in \eqref{DHKM:prodwts}  one more time. We note that
$\bignorm[p]{\bslambda_d^{(i_0)}}^p<\infty$ holds if and only if  $\bignorm[p]{\bslambda_d}^p<\infty$.
We denote the $d$-dependence of the vector of coordinate weights by writing $\bsw_d:=(w_1, \ldots, w_d)$ and $\bsw_\infty = (w_1, w_2, \ldots)$
By Lemma \ref{DHKM:prodwtsupperbound}, we have
\[
\norm[p]{\bss} \norm[p]{\bsw_d} \le \norm[p]{\bslambda_d}\le \exp\left(\frac 1p \norm[p]{\bss}^p \norm[p]{\bsw_d}^p  \right),
\]
and this estimate holds for any $d\in\NN$.
Thus, we have necessary and sufficient conditions for strong tractability.
\begin{corollary} \label{DHKM:sptexample_cor}
For the $\lambda_{d,\bsk}$ of the form \eqref{DHKM:prodwts} we have strong polynomial tractability if and only if there exists $p>0$ such that
 \begin{equation*}
%\label{DHKM:sptexample}
    \norm[p]{\bss} < \infty \text{ and } 
    \norm[p]{\bsw_\infty} < \infty.
\end{equation*}
In particular, considering example in Section \ref{DHKM:secexamp}, where $s_{k} = k^{-r}$, we have strong polynomial tractability if and only if there exists $p> 0 $ such that
\begin{equation*}
%\label{DHKM:sptexample}
   r p > 1 \text{ and } 
    \norm[p]{\bsw_\infty} < \infty.
\end{equation*}
\end{corollary}

@@

Let $\DHKMchange{\bsn = (n_j)_{j\ge 0}}$ be a strictly increasing sequence of non-negative integers.  This sequence may increase geometrically or algebraically. Define the sets of wavenumbers analogously to \eqref{DHKM:KOnedef},
\begin{equation*}
   n_{-1}=0, \qquad \calK_j := \{\bsk_{n_{j-1}+1}, \ldots, \bsk_{n_j}\} \quad \text{for } j \in \N_0.
\end{equation*}
If $n_0 = 0$, then $\calK_0$ is empty.  For any $f \in \calF$, define the norms of subsets of series coefficients:
\begin{equation} \label{DHKM:SigmaDef}
\sigma_j (f):=\norm[\rho]{\biggl(\frac{\hf(\bsk)}{\lambda_{\bsk}} \biggr)_{\bsk \in \calK_j}} \qquad \text{for } j \in \N.
\end{equation}
Thus, $\norm[\calF]{f} = \norm[\rho]{\bigl(\sigma_j(f) \bigr)_{j \in \N_0}}$.

\DHKMchange{As stated above, we would like to keep track of the rate of decay of the $\hf(\bsk_i)$. Hence, for this section, we define the cone of input functions by}
\begin{equation} \label{DHKM:TrackConeDef}
  \DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}} : =\left\{f\in\calF \colon \sigma_{j+s} (f)\le ab^s \sigma_j (f)\ \forall j,s\in\NN\right\}.
\end{equation}
Here, $a$ and $b$ are positive reals with $b< 1 < a$. The constant $a$ is an inflation factor, and the constant $b$ defines the general rate of decay of the $\sigma_j(f)$ for $f \in  \DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}} $. 
\DHKMchange{We remark that knowing $\sigma_1 (f)$ would already be sufficient to identify 
a sample size that guarantees an error of at most $\varepsilon$. However, knowing $\sigma_j (f)$ for further values of $j$ will allow us to stop earlier.}

Because $ab^s$ may be greater than one, we do not require the series coefficients of the solution, $\SOL(f)$, to decay monotonically. However, we expect their partial sums to decay steadily.  The series coefficients for wavenumbers $\bsk \in \calK_0$ do not affect the definition of $\DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}} $ and may behave erratically.
Lemma \ref{DHKM:Key_Lem} implies that 
\begin{equation} \label{DHKM:LambdaDef}
    \norm[\tau]{\bigl(\hf(\bsk) \bigr)_{\bsk \in \calK_j}} \le \sigma_j(f) \Lambda_j, \qquad \text{where } \Lambda_j : = \norm[\rho']{\bigl(\lambda_{\bsk} \bigr)_{\bsk \in \calK_j}}.
\end{equation}
From \eqref{DHKM:SOLNorm} and \eqref{DHKM:SOLNormFinite} it follows that the norm of the solution operator is 
\begin{equation} \label{DHKM:NormLambdaFinite}
    \norm[\calF \to \calG]{\SOL} = \bignorm[\rho']{\bigl(\Lambda_j \bigr)_{j \in \N_0}} < \infty.
\end{equation}

If $f$ belongs to the $\DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}} $ defined in \eqref{DHKM:TrackConeDef} and $n_0 = 0$, then 
\begin{align*}
    \norm[\calF]{f} & = \norm[\rho]{\Biggl(\,\norm[\rho]{\biggl(\frac{\hf(\bsk)}{\lambda_{\bsk}} \biggr)_{\bsk \in \calK_j}}\,\Biggr)_{j \in \N}} = \bignorm[\rho]{\bigl(\sigma_j(f) \bigr)_{j \in \N}} \\
    & \le \bignorm[\rho]{\bigl(\sigma_1(f), ab \sigma_1(f), ab^2 \sigma_1(f), \ldots \bigr)} \\
    & = \left(1 + \frac{a^\rho b^\rho}{1 - b^\rho} \right)^{1/\rho}  \norm[\rho]{\biggl(\frac{\hf(\bsk)}{\lambda_{\bsk}} \biggr)_{\bsk \in \calK_1}}.
\end{align*}
Comparing this inequality to the definition of $\DHKMchange{\calC_{\boldsymbol{\lambda},n_1,A}} $ in the previous section, it can be seen that $\DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}} $  defined in \eqref{DHKM:TrackConeDef} is a subset of  $\DHKMchange{\calC_{\boldsymbol{\lambda},n_1,A}}$ defined in \eqref{DHKM:pilot_cone} if we choose 
$A=\left[1 + a^\rho b^\rho/(1 - b^\rho) \right]^{1/\rho}$ in \eqref{DHKM:pilot_cone}.

From the expression for the error in \eqref{DHKM:APP_Err_Coef} and the definition of the cone in  \eqref{DHKM:TrackConeDef}, we can now derive a data-driven error bound for all $f \in \DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}} $ and $j \in \bbN$: 
\begin{align}
\nonumber
\MoveEqLeft{\norm[\calG]{\SOL(f)-\APP(f,n_j)}} \\
\nonumber &= \norm[\tau]{\left(\hf(\bsk_i) \right)_{i = n_j+1}^\infty}
= \norm[\tau]{ \left(\norm[\tau]{\bigl(\hf(\bsk) \bigr)_{\bsk \in \calK_l}} \right)_{l=j+1}^\infty}
\\
\nonumber
& \le \norm[\tau]{ \bigl(\sigma_l(f) \Lambda_l \bigr)_{l=j+1}^\infty} \qquad \text{by \eqref{DHKM:LambdaDef}} \\
\nonumber 
&
= \norm[\tau]{ \bigl(\sigma_{j+s}(f) \Lambda_{j+s} \bigr)_{s=1}^\infty}
\\
& \le a \sigma_j(f) \norm[\tau]{ \bigl(b^s\Lambda_{j+s} \bigr)_{s=1}^\infty} =:\ERRNj
 \qquad \text{by \eqref{DHKM:TrackConeDef}.}
 \label{DHKM:algoineq}
\end{align}

@@

The data-driven error bound in \eqref{DHKM:algoineq} forms the basis for an adaptive Algorithm \ref{DHKM:TrackConeAlg}, which solves our problem for input functions in the cone $\DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}}$ defined in \eqref{DHKM:TrackConeDef}.  The following theorem establishes its viability and computational cost. In deriving upper bounds on the computational cost and lower bounds on the complexity, we may sacrifice tightness for simplicity.

@@

The following theorem establishes a lower bound on the complexity of our problem for input functions in $\DHKMchange{\calC_{\boldsymbol{\lambda},\bsn,a,b}}$. The theorem after that shows that the cost of our algorithm as given in Theorem \ref{DHKM:TractConeCompCost} is essentially no worse than this lower bound.
